Predicting Political Support in the United States

### 2024-03-17 output/paper.qmd
Properly cite

The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump \[cite2020Election\]. Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors. We will be using a representative sample of 60000 Americans to discern patterns in these voters' responses and how their personal backgrounds may have influenced their voting decision. We will be examining the impact of various factors such as: gender, race, educational background, religion, and personal stances on key political issues such as gun control and student loans. Our main focus is to compare the correlations in the 2022 CES survey data to the 2020 CES survey data, and to see if our estimand, the effect of a voter's cultural background and political stances on their voting decision, has influenced their voting decision in the 2020 election. Using these understandings, we hope to highlight the potential implications these correlations may have for this upcoming 2024 election.

@@

Although there seems to be a strong correlation between if a voter identifies as an other gender to supporting Biden, and a strong correlation between if a voter identifies as an other race to supporting Biden, but these must be considered carefully as within the sample (with seed 302), only 5 and 1 voters identified in those categories respectively. There were various other demographics in the dataset that had relatively small sample sizes, such as: `race` where non black, white, or hispanic respondents made up a small portion of the sample, or non Catholic, Protestant, or Atheist respondents in `religion`. To get a more accurate reading in regards to these features, it would be necessary to train a model which includes a larger sample size of these demographics or to find a way to account for these values while still making use of them. This was a scenario that one may expect to see as there can always be underrepresented demographics in a 60000 respondent sized sample. This is a big limitation to our models, and is something that should be considered when interpreting the results.

>>>>

The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump @cite2020Election. Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors. We will be using a representative sample of 60000 Americans to discern patterns in these voters' responses and how their personal backgrounds may have influenced their voting decision. We will be examining the impact of various factors such as: gender, race, educational background, religion, and personal stances on key political issues such as gun control and student loans. Our main focus is to compare the correlations in the 2022 CES survey data to the 2020 CES survey data, and to see if our estimand, the effect of a voter's cultural background and political stances on their voting decision, has influenced their voting decision in the 2020 election. Using these understandings, we hope to highlight the potential implications these correlations may have for this upcoming 2024 election.

@@

Although there seems to be a strong correlation between if a voter identifies as an other gender to supporting Biden, and a strong correlation between if a voter identifies as an other race to supporting Biden, but these must be considered carefully as within the sample (with seed 302), only 5 and 1 voters identified in those categories respectively. There were various other demographics in the dataset that had relatively small sample sizes, such as: `race` where non Black, White, or Hispanic respondents made up a small portion of the sample, or non Catholic, Protestant, or Atheist respondents in `religion`. To get a more accurate reading in regards to these features, it would be necessary to train a model which includes a larger sample size of these demographics or to find a way to account for these values while still making use of them. This was a scenario that one may expect to see as there can always be underrepresented demographics in a 60000 respondent sized sample. This is a big limitation to our models, and is something that should be considered when interpreting the results.
